### 2021-05-28 thesis/tex/sterileTheory.tex
update

\documentclass{article}
\input{commands}
\usepackage{physics,amsmath, amsfonts, siunitx, amssymb, subcaption}
\usepackage[utf8]{inputenc}
\usepackage[margin=0.5in]{geometry}
\usepackage{xr-hyper}
%\externaldocument{sterile}

\begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}. The anomaly was consistent with a fourth
neutrino state with $\dm[41] > \SI{0.03}{\eV \squared}$. Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly, but saw the excess in The
$\ne$ events too. Together with the so-called reactor and gallium analomalies, there were indications that both appearance and disappearance
anomalies might be remedied by a fourth mass state.

@@

The inclusion of the sterile neutrino in the Hamiltonian is straightforward. We extend the PMNS matrix to incorporate the new flavor and mass eigenstates: 
\begin{align}
    U_{4gen} =
    \begin{pmatrix}
    U_{e1} & U_{e2} & U_{e3} & U_{e4} \\
    U_{\mu1} & U_{\mu2} & U_{\mu3} & U_{\mu4} \\
    U_{\tau1} & U_{\tau2} & U_{\tau3} & U_{\tau4} \\
    U_{s1} & U_{s2} & U_{s3} & U_{s4}
    \end{pmatrix}\,,
\end{align}
where a common parametrization of the PMNS matrix is 
\begin{align}\label{eq:U4gen_param}
    U_{4gen} = R_{34}R_{24}R_{14}U_{3gen}\,.
\end{align}
The mass matrix extends analogously:
\begin{align}\label{eq:dm4gen}
    M^2_{4gen} =
    \begin{pmatrix}
        0 & 0 & 0 & 0\\
        0 & \dm_{21} & 0  & 0 \\
        0 & 0 & \dm_{31} & 0 \\
        0 & 0 & 0 & \dm_{41} \\
    \end{pmatrix}\,.
\end{align}

Now,the interaction with matter requires a careful reconsideration of the matter potential. We start of with the unaltered potential matrix. 
Just as with the PMNS matrix, we extend this to $4\times4$:
\begin{align}\label{eq:Vsterile1}
    \begin{pmatrix}
        V_{CC} + V_{NC} & 0 & 0 & 0 \\
        0 &V_{NC} & 0 & 0 \\
        0 & 0 & V_{NC} & 0 \\
        0 & 0 & 0 & 0 
    \end{pmatrix}\,.
\end{align}
Now we need to include the terms that describes the matter potential felt by the sterile flavor state. Recalling our discussion above, we remind ourselves that the sterile neutrino by definition
does not participate in any interaction\footnote{The exception to this is of course gravity. The sterile neutrino is not massless.}. Thus, all potential terms involving the sterile state are zero. 
In other words, the potential matrix in Eq.~\ref{eq:Vsterile1} is complete, save for the usual subtraction by a constant identity matrix:

@@

\section{Sterile Signals}
Since the new neutrino does not interact weakly, how do we then detect its signal? If the sterile mixing angle $\theta_{i4}$ is non-zero, 
we allow the sterile mass state to mix with the active state $i$. The most interesting case is when $\theta_{24} \neq 0$, 
which for $\dm[41] \sim \si{\eV \squared}$ gives rise to a resonant disappeareance in the $\bar\mu \bar\mu$ sector, shown in Fig.~\ref{fig:sterile_resonance}.

\begin{figure}
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=1\linewidth]{figures/sterile_resonance.pdf}
        \caption{$\Pamam$ disappearance}\label{fig:sterile_resonance}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=1\linewidth]{figures/resonance_shift.pdf}
        \caption{$\Pamam$ disappearance}\label{fig:resonance_shift}
    \end{subfigure}
\end{figure}

So for a \si{\eV}-scale sterile neutrino and a non-zero $\theta_{24}$, we expect a \si{\TeV} $\anm$ disappearance.
The resonance dip is affected by the value of $\dm[41]$ and $\theta_{24}$ as shown in Fig.~\ref{fig:resonance_shift}.
We see how the value of $\dm[41]$ shifts the peak, while the mixing angle adjusts its strength.

\newpage
\bibliographystyle{elsarticle-num}
\bibliography{ref}
\end{document}

>>>>

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{subcaption}
% \input{commands}

% \begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}. The anomaly was consistent with a fourth
neutrino state with $\dm[41] > \SI{0.03}{\eV \squared}$. Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly, but saw the excess in The
$\ne$ events too. Together with the so-called reactor and gallium analomalies, there were indications that both appearance and disappearance
anomalies might be remedied by a fourth mass state.

@@

\subsection{The Hamiltonian}
The inclusion of the sterile neutrino in the Hamiltonian is straightforward. We extend the PMNS matrix to incorporate the new flavor and mass eigenstates: 
\begin{align}
    U_{4gen} =
    \begin{pmatrix}
    U_{e1} & U_{e2} & U_{e3} & U_{e4} \\
    U_{\mu1} & U_{\mu2} & U_{\mu3} & U_{\mu4} \\
    U_{\tau1} & U_{\tau2} & U_{\tau3} & U_{\tau4} \\
    U_{s1} & U_{s2} & U_{s3} & U_{s4}
    \end{pmatrix}\,,
\end{align}
where a common parametrization of the PMNS matrix is 
\begin{align}\label{eq:U4gen_param}
    U_{4gen} = R_{34}R_{24}R_{14}U_{3gen}\,.
\end{align}
The mass matrix extends analogously:
\begin{align}\label{eq:dm4gen}
    M^2_{4gen} =
    \begin{pmatrix}
        0 & 0 & 0 & 0\\
        0 & \dm_{21} & 0  & 0 \\
        0 & 0 & \dm_{31} & 0 \\
        0 & 0 & 0 & \dm_{41} \\
    \end{pmatrix}\,.
\end{align}

Now, the interaction with matter requires a careful reconsideration of the matter potential. We start of with the unaltered potential matrix. 
Just as with the PMNS matrix, we extend this to $4\times4$:
\begin{align}\label{eq:Vsterile1}
    \begin{pmatrix}
        V_{CC} + V_{NC} & 0 & 0 & 0 \\
        0 &V_{NC} & 0 & 0 \\
        0 & 0 & V_{NC} & 0 \\
        0 & 0 & 0 & 0 
    \end{pmatrix}\,.
\end{align}
Now we need to include the terms that describes the matter potential felt by the sterile flavor state. Recalling our discussion above, we remind ourselves that the sterile neutrino by definition
does not participate in any interaction\footnote{The exception to this is of course gravity. The sterile neutrino is not massless.}. Thus, all potential terms involving the sterile state are zero. 
In other words, the potential matrix in Eq.~\ref{eq:Vsterile1} is complete, save for the usual subtraction by a constant identity matrix:

@@

\subsection{Sterile Signals}
Since the new neutrino does not interact weakly, how do we then detect its signal? If the sterile mixing angle $\theta_{i4}$ is non-zero, 
we allow the sterile mass state to mix with the active state $i$. The most interesting case is when $\theta_{24} \neq 0$, 
which for $\dm[41] \sim \si{\eV \squared}$ gives rise to a resonant disappeareance in the $\bar\mu \bar\mu$ sector, shown in Fig.~\ref{fig:sterile_resonance}.

\begin{figure}
    \centering
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=1\linewidth]{figures/sterile_resonance.pdf}
        \caption{$\Pamam$ disappearance}\label{fig:sterile_resonance}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \includegraphics[width=1\linewidth]{figures/resonance_shift.pdf}
        \caption{$\Pamam$ disappearance}\label{fig:resonance_shift}
    \end{subfigure}
\end{figure} %TODO: better captions

So for a \si{\eV}-scale sterile neutrino and a non-zero $\theta_{24}$, we expect a \si{\TeV} $\anm$ disappearance.
The resonance dip is affected by the value of $\dm[41]$ and $\theta_{24}$ as shown in Fig.~\ref{fig:resonance_shift}.
We see how the value of $\dm[41]$ shifts the peak, while the mixing angle adjusts its strength. 
%TODO: more? mention that Pmm=1 at TeV for 3+0

% \newpage
% \bibliographystyle{elsarticle-num}
% \bibliography{ref}
% \end{document}
